## R script for sea trutle stable isotope analysis.

Cyler Conrad, Department of Anthropology, University of New Mexico, cylerc@unm.edu

This document contain R code to reproduce the plots and statistical analysis presented in: 

> Conrad, C., L. Pagès Barceló, J.A. Seminoff, C.N. Turner Tomaszewicz, M.J. Labonte, B.M. Kemp, E.L. Jones, S.D. Newsome, M. Stoyka, K. Bruner and A.G. Pastron. in preparation. Stable Isotope Ecology and Ancient DNA Analysis of Sea Turtles (Cheloniidae) from the Gold Rush-era (1850s) Eastern Pacific Ocean. Open Quaternary. 

All data required to perform the analyses can be found at the University of New Mexico digital electronic repository (Conrad et al. 2017). The development version of this document can be found at https://github.com/cylerc/OQ_SeaTurtle

Details of the computational environment and software dependencies are listed at the end of this document. 

Reference: Conrad, C., L. Pagès Barceló, J.A. Seminoff, C.N. Turner Tomaszewicz, M.J. Labonte, B.M. Kemp, E.L. Jones, S.D. Newsome, M. Stoyka, K. Bruner and A.G. Pastron. 2017. Historic and modern eastern Pacific green sea turtle stable isotopes [dataset]. University of New Mexico.

```{r setup}
# set the base directory for knitr to the directory above this one
# install.packages(c("ggplot2"))
library(knitr)
opts_knit$set(root.dir = '../', progress = FALSE)
```

```{r load_libraries, message = FALSE, warning = FALSE}
# see the output of sessionInfo() at the bottom for package version numbers
library(ggplot2)
library(ggmap)
```

```{r figure 2-map of sites}
#load coordinate data
sites<-read.csv("coords.csv")
region<-qmap("Baja California", zoom=5)
region + geom_point(aes(x = long, y = lat, shape=Sample), data = sites)
```

```{r figure 3-bone to bone isotopes and statistics}
#load stable isotope data
master<-read.csv("data/master.csv",stringsAsFactors=FALSE)
#extract data for figure 3 plot and analysis
fig3<-master[1:17,]
#plot the results
ggplot(fig3, aes(d13C, d15N, shape=Site)) +
  geom_point(size=5) +
  xlab(expression(paste(delta^{13},'C'))) +
  ylab(expression(paste(delta^{15},'N'))) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(colour="white", fill="white"),
        strip.text.x = element_text(colour = "black", size = 15),
        axis.title.x = element_text(colour="black", size=25),
        axis.text.x  = element_text(colour="black", vjust=0.5, size=20),
        axis.title.y = element_text(face="bold", colour="black", size=25),
        axis.text.y  = element_text(colour="black", vjust=0.5, size=20),
        legend.title = element_text(color="black", size=16),
        legend.position="right",
        legend.text = element_text(color="black", size=15)) +
  scale_shape_manual(values=c(1,0,15))+
  theme(legend.position="right")
#break data into groups for statistical analysis
gr<-fig3[1:2,]
modern<-fig3[3:17,]
wilcox.test(gr$d13C, modern$d13C, conf.int = TRUE)
wilcox.test(gr$d15N, modern$d15N, conf.int = TRUE)
mean(modern$d13C)
sd(modern$d13C)
```

```{r figure 4-bone to skin isotopes and statistics}
#extract data for figure 4 plot and analysis
fig4<-master[18:26,]
fig4$sd <- as.numeric(as.character(fig4$sd))
fig4$sd.1 <- as.numeric(as.character(fig4$sd.1))
#plot the results
ggplot(fig4, aes(x=d13C, y=d15N, shape=Site)) +
  geom_point(size=5) +    
  geom_errorbarh(aes(xmin=d13C-sd,xmax=d13C+sd),height=0.2)+
  geom_errorbar(aes(ymin=d15N-sd.1,ymax=d15N+sd.1),width=0.2) +
  xlab(expression(paste(delta^{13},'C')))+
  ylab(expression(paste(delta^{15},'N')))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        strip.background = element_rect(colour="white", fill="white"),
        strip.text.x = element_text(colour = "black", size = 15),
        legend.title = element_text(colour="black", size=15, face="bold"),
        legend.text = element_text(size=14),
        axis.text.x = element_text(vjust=0.5, color="black", size=15, face="bold"),
        axis.text.y = element_text(vjust=0.5, color="black", size=15, face="bold"), 
        axis.title.y = element_text(vjust=1.0, color="black", size=18, face="bold"),
        axis.title.x = element_text(vjust=0.1, color="black", size=18, face="bold"),
        plot.title = element_text(color="black", size=15, face="bold")) +
  scale_shape_manual(values=c(3,15,16,17,0,1))+
  theme(legend.position="right")
#break data into groups for statistical analysis
modern2<-fig4[1:7,]
gr2<-fig4[8:9,]
wilcox.test(gr2$d13C, modern2$d13C, conf.int = TRUE)
wilcox.test(gr2$d15N, modern2$d15N, conf.int = TRUE)
mean(modern2$d13C)
sd(modern2$d13C)
```

```{r session_info}
# Computational environment and software dependencies for this analysis.
sessionInfo()
```

LICENSE
 
The MIT License (MIT)

Copyright (c) 2017 Cyler Conrad

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in
all copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
THE SOFTWARE.
